---
title: "SQL"
author: "Adam Craddock"
date: "December 14, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Filtering Columns & Rows

In SQL you can filter columns and rows using SELECT and WHERE to find rows. Using the Lahman baseball database, we will provide an example.

First, download the needed packages written below.


```{r}
library(Lahman)
library(sqldf)

```

Next, the example, suppose we wanted to see the total HR totals for the 1927 Yankees. The code for this would be:

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting WHERE teamID='NYA' and yearID=1927"

sqldf(query)


```

Getting a bit more specific, this is extracting which NY Yankees hit 40 or more HRs
```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting WHERE teamID='NYA' and HR>=40"
sqldf(query)
```

##Strike Outs (SO) & Homeruns(HR)
More filtering allows extraction of who had over 40 HRs and less than 60 SOs

```{r}
query<-"SELECT playerID,teamID,yearID,HR,SO FROM Batting WHERE HR>40 and SO<60"
sqldf(query)
```

#Filtering the Teams
Finally for this section, we can filter out the teams to just the Phillies and finding all players in the 1970s who had over 30 HRs

One filters out from the Batting table for the Phillies and the years between 1970 and 1979 for Homeruns greater than 30. 

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting WHERE teamID='PHI' and yearID>=1970 and yearID<=1979 and HR>30"
sqldf(query)
```


```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)
```


```{r}
query<-"SELECT playerID,yearID,teamID,SO,AB FROM Batting
WHERE SO<10 and AB>=400 
ORDER BY SO"

sqldf(query)
```


```{r}
query<-"SELECT HR FROM Batting
WHERE playerID= 'ruthba01'"

sqldf(query)
```


```{r}
query<-"SELECT playerID, sum (HR) FROM Batting
WHERE playerID='ruthba01'
GROUP BY playerID"

sqldf(query)
```

Grouping data by playerID with conditions such as average Homeruns being over 600 using the function HAVING then ORDER BY average Homeruns in descending 

```{r}
query<-"SELECT playerId, sum(HR) FROM Batting
GROUP BY playerID
HAVING sum(HR)>600
ORDER BY sum(HR) DESC"


sqldf(query)
```


Isolating the First Name, Last Name, and Homeruns for Babe Ruth, one of the greats.

```{r}
query<-"SELECT nameFirst,nameLast,yearID,teamID,HR 
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.playerID='ruthba01'"

sqldf(query)
```

Here we are joining lists again using the INNER JOIN Function as well as ordering the data by Homeruns in descending order.

```{r}
query<-"SELECT nameFirst,nameLast,yearID,teamID,HR
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)
```

One can isolate the statistics of Babe Ruth through joining various lists and pulling his name from the batting and teams table.
```{r}
query<-"SELECT playerID,name,Batting.yearID,Batting.HR
FROM Batting INNER JOIN Teams
ON Batting.teamID=Teams.teamID AND Batting.yearID=Teams.yearID
WHERE playerID='ruthba01'"

sqldf(query)
```


#Joining lists and columns allows for the analyst to find the first and last names of all players named 'Bob' who had an average salary of over 1,000,000 USD

```{r}
query<-"SELECT nameFirst,nameLast,avg(salary)
FROM Salaries INNER JOIN Master
ON Salaries.playerId=Master.playerID
WHERE nameFirst='Bob'
GROUP BY Salaries.playerID
HAVING avg(salary)>1000000
ORDER BY avg(salary) DESC"

sqldf(query)
```

```

